---
title: "Assessment Intelligence Modeling Report"
subtitle: "Synthetic assessment data generated for a static portfolio dashboard"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Purpose

This report documents the analysis layer behind the public assessment dashboard.
The dashboard itself is static JavaScript, but the data build and statistical
summaries are produced ahead of time in R.

The source data used here is synthetic. It is designed to preserve dashboard
behavior and statistical structure without publishing real students, rosters,
grades, LMS records, or school-private exports.

## Load Data

```{r load-data}
student_records <- read.csv("../data/synthetic/student_level_assessment.csv", stringsAsFactors = FALSE)
section_summary <- read.csv("../data/synthetic/section_period_summary.csv", stringsAsFactors = FALSE)

str(student_records)
str(section_summary)
```

## Program Trend

```{r program-trend}
weighted_mean <- function(values, weights) {
  if (!sum(weights)) return(0)
  sum(values * weights) / sum(weights)
}

period_summary <- do.call(
  rbind,
  lapply(split(section_summary, section_summary$order), function(rows) {
    data.frame(
      periodLabel = rows$periodLabel[[1]],
      order = rows$order[[1]],
      score = weighted_mean(rows$score, rows$students),
      completion = weighted_mean(rows$completion, rows$students),
      proficiency = weighted_mean(rows$proficiency, rows$students)
    )
  })
)

period_summary
```

```{r trend-plot}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(period_summary, ggplot2::aes(order, score)) +
    ggplot2::geom_line(color = "#003260", linewidth = 1) +
    ggplot2::geom_point(color = "#c69a36", size = 2.5) +
    ggplot2::scale_x_continuous(
      breaks = period_summary$order,
      labels = period_summary$periodLabel
    ) +
    ggplot2::labs(
      title = "Synthetic Assessment Score Trend",
      x = NULL,
      y = "Weighted mean score"
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 35, hjust = 1))
} else {
  plot(period_summary$order, period_summary$score, type = "b",
       xlab = "Assessment period", ylab = "Weighted mean score",
       main = "Synthetic Assessment Score Trend")
}
```

## Growth Model

```{r growth-model}
section_summary$season <- factor(section_summary$season, levels = c("Fall", "Spring"))
section_summary$course <- factor(section_summary$course)
section_summary$order_centered <- section_summary$order - min(section_summary$order)

growth_model <- lm(
  score ~ order_centered + season + course + completion,
  data = section_summary,
  weights = pmax(section_summary$completed, 1)
)

summary(growth_model)
```

## Completion Model

```{r completion-model}
completion_model <- glm(
  cbind(completed, notCompleted) ~ order_centered + season + course,
  data = section_summary,
  family = binomial()
)

summary(completion_model)
```

## Interpretation Notes

- The models are portfolio artifacts, not claims about a real institution.
- The growth model demonstrates how section/course/time effects can be modeled
  before exporting dashboard-ready summaries.
- The completion model separates participation behavior from performance.
- In a private deployment, the same structure could support real internal
  analysis while publishing only synthetic or aggregate-safe outputs.

