---
title: "Assessment Intelligence Modeling Report"
subtitle: "Synthetic assessment data generated for a static portfolio dashboard"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Purpose

This report documents the analysis layer behind the public assessment dashboard.
The dashboard itself is static JavaScript, but the data build and statistical
summaries are produced ahead of time in R.

The source data used here is synthetic. It is designed to preserve dashboard
behavior and statistical structure without publishing real students, rosters,
grades, LMS records, or school-private exports.

## Recommendation

Use the synthetic dashboard as a public demonstration of assessment intelligence:
the R build layer creates fake but realistic assessment records, models growth
and completion patterns, and exports dashboard-ready JSON for a static web
frontend.

The report should be read as a reproducible workflow artifact, not as evidence
about a real school.

## Direct Answers

- **What powers the dashboard?** A synthetic assessment dataset generated and
  summarized before deployment.
- **What does the model layer add?** It separates growth, completion, and course
  effects so the dashboard can support decision conversations rather than only
  display raw scores.
- **What is public-safe?** The records use fake students, fake teachers, fake
  sections, synthetic scores, and validation checks.

## Load Data

```{r load-data}
student_records <- read.csv("../data/synthetic/student_level_assessment.csv", stringsAsFactors = FALSE)
section_summary <- read.csv("../data/synthetic/section_period_summary.csv", stringsAsFactors = FALSE)

str(student_records)
str(section_summary)
```

## Data Audit

```{r data-audit}
data.frame(
  Metric = c(
    "Student-period records",
    "Section-period records",
    "Synthetic students",
    "Sections",
    "Courses",
    "Assessment periods"
  ),
  Value = c(
    nrow(student_records),
    nrow(section_summary),
    length(unique(student_records$studentId)),
    length(unique(section_summary$sectionId)),
    length(unique(section_summary$course)),
    length(unique(section_summary$periodId))
  )
)
```

## Program Trend

```{r program-trend}
weighted_mean <- function(values, weights) {
  if (!sum(weights)) return(0)
  sum(values * weights) / sum(weights)
}

period_summary <- do.call(
  rbind,
  lapply(split(section_summary, section_summary$order), function(rows) {
    data.frame(
      periodLabel = rows$periodLabel[[1]],
      order = rows$order[[1]],
      score = weighted_mean(rows$score, rows$students),
      completion = weighted_mean(rows$completion, rows$students),
      proficiency = weighted_mean(rows$proficiency, rows$students)
    )
  })
)

period_summary
```

```{r trend-plot}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(period_summary, ggplot2::aes(order, score)) +
    ggplot2::geom_line(color = "#003260", linewidth = 1) +
    ggplot2::geom_point(color = "#c69a36", size = 2.5) +
    ggplot2::scale_x_continuous(
      breaks = period_summary$order,
      labels = period_summary$periodLabel
    ) +
    ggplot2::labs(
      title = "Synthetic Assessment Score Trend",
      x = NULL,
      y = "Weighted mean score"
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 35, hjust = 1))
} else {
  plot(period_summary$order, period_summary$score, type = "b",
       xlab = "Assessment period", ylab = "Weighted mean score",
       main = "Synthetic Assessment Score Trend")
}
```

## Model Journey

The first modeling pass is intentionally simple and inspectable. It uses:

- weighted linear modeling for score growth
- binomial modeling for completion
- course and season terms to separate structural effects
- synthetic group sizes as weights where appropriate

Future iterations can add mixed models, splines, or tree-based examples after
the public-safe workflow is stable.

## Final Growth Model

```{r growth-model}
section_summary$season <- factor(section_summary$season, levels = c("Fall", "Spring"))
section_summary$course <- factor(section_summary$course)
section_summary$order_centered <- section_summary$order - min(section_summary$order)

growth_model <- lm(
  score ~ order_centered + season + course + completion,
  data = section_summary,
  weights = pmax(section_summary$completed, 1)
)

summary(growth_model)
```

## Completion Model

```{r completion-model}
completion_model <- glm(
  cbind(completed, notCompleted) ~ order_centered + season + course,
  data = section_summary,
  family = binomial()
)

summary(completion_model)
```

## Metric Scale

```{r metric-scale}
latest_period <- max(section_summary$order)
latest_rows <- section_summary[section_summary$order == latest_period, ]
data.frame(
  Metric = c("Weighted score", "Weighted completion", "Weighted proficiency"),
  Value = c(
    weighted_mean(latest_rows$score, latest_rows$students),
    weighted_mean(latest_rows$completion, latest_rows$students),
    weighted_mean(latest_rows$proficiency, latest_rows$students)
  )
)
```

## Model Checks

```{r model-checks}
growth_residuals <- residuals(growth_model)
data.frame(
  Check = c("Mean residual", "Median residual", "Residual SD"),
  Value = c(
    mean(growth_residuals),
    median(growth_residuals),
    sd(growth_residuals)
  )
)
```

## Sensitivity Check

```{r sensitivity-check}
target_values <- c(65, 70, 75)
data.frame(
  ProficiencyTarget = target_values,
  LatestPercentAtOrAboveTarget = sapply(target_values, function(target) {
    mean(student_records$score[
      student_records$order == latest_period & student_records$completed
    ] >= target) * 100
  })
)
```

## Decision Scenarios

- If completion is below target, interpret performance trends cautiously and
  inspect participation first.
- If growth is positive but skill signals remain uneven, use the report to pick
  reteaching priorities rather than declare broad success.
- If course-level trends diverge, review pacing, assessment alignment, and
  intervention practices before making personnel conclusions.

## Interpretation Notes

- The models are portfolio artifacts, not claims about a real institution.
- The growth model demonstrates how section/course/time effects can be modeled
  before exporting dashboard-ready summaries.
- The completion model separates participation behavior from performance.
- In a private deployment, the same structure could support real internal
  analysis while publishing only synthetic or aggregate-safe outputs.

## Bottom Line

The reporting artifact should give leaders an answer, show enough model context
to trust the workflow, name the caveats, and end with a practical next action.
