---
title: "Synthetic Gradebook Reconstruction Report"
subtitle: "R-generated public-safe student score dataset"
output:
  html_document:
    toc: true
    toc_depth: 2
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Purpose

This report summarizes the public-safe synthetic gradebook reconstruction
workflow. The source artifacts are generated by R from private reference
structure, but this report reads only synthetic outputs.

## Load Synthetic Outputs

```{r load-data}
wide_gradebook <- read.csv("data/synthetic/synthetic_gradebook.csv", stringsAsFactors = FALSE, check.names = FALSE)
score_records <- read.csv("data/synthetic/synthetic_student_scores_long.csv", stringsAsFactors = FALSE)
assignment_metadata <- read.csv("data/synthetic/synthetic_assignment_metadata.csv", stringsAsFactors = FALSE)

str(wide_gradebook)
str(score_records)
str(assignment_metadata)
```

## Output Audit

```{r output-audit}
data.frame(
  Metric = c(
    "Wide gradebook rows",
    "Wide gradebook columns",
    "Long-form score records",
    "Synthetic students",
    "Synthetic sections",
    "Assignment metadata rows",
    "Assignment families",
    "Skill domains"
  ),
  Value = c(
    nrow(wide_gradebook),
    ncol(wide_gradebook),
    nrow(score_records),
    length(unique(score_records$synthetic_student_id)),
    length(unique(score_records$synthetic_section)),
    nrow(assignment_metadata),
    length(unique(score_records$assignment_family)),
    length(unique(score_records$skill_domain))
  )
)
```

## Completion And Missingness

```{r completion-summary}
completion_summary <- aggregate(
  completed ~ assignment_family,
  data = score_records,
  FUN = function(x) round(100 * mean(x), 1)
)
names(completion_summary)[names(completion_summary) == "completed"] <- "completion_rate"
completion_summary
```

```{r missingness-summary}
missingness_summary <- as.data.frame(table(score_records$missingness_reason))
names(missingness_summary) <- c("missingness_reason", "records")
missingness_summary[order(-missingness_summary$records), ]
```

## Score Distribution

```{r score-distribution}
completed_scores <- score_records[score_records$completed & !is.na(score_records$score), ]
if (any(!is.na(completed_scores$score_percent))) {
  completed_scores$score_metric <- completed_scores$score_percent
  score_metric_label <- "Score percent"
} else {
  completed_scores$score_metric <- completed_scores$score
  score_metric_label <- "Synthetic score"
}

data.frame(
  Metric = c("Completed score records", paste("Mean", score_metric_label), paste("Median", score_metric_label), paste(score_metric_label, "SD")),
  Value = c(
    nrow(completed_scores),
    round(mean(completed_scores$score_metric), 1),
    round(median(completed_scores$score_metric), 1),
    round(sd(completed_scores$score_metric), 1)
  )
)
```

```{r score-plot}
if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(completed_scores, ggplot2::aes(score_metric)) +
    ggplot2::geom_histogram(binwidth = 5, fill = "#2563eb", color = "white") +
    ggplot2::facet_wrap(~ assignment_family) +
    ggplot2::labs(
      title = "Synthetic Score Distribution By Assignment Family",
      x = score_metric_label,
      y = "Completed records"
    ) +
    ggplot2::theme_minimal()
} else {
  hist(completed_scores$score_metric, breaks = 20, main = "Synthetic Score Distribution", xlab = score_metric_label)
}
```

## Latent Pattern Checks

```{r latent-patterns}
aggregate(
  score_metric ~ ability_band + engagement_band,
  data = completed_scores,
  FUN = function(x) round(mean(x), 1)
)
```

```{r completion_by_risk_band}
aggregate(
  completed ~ risk_band,
  data = score_records,
  FUN = function(x) round(100 * mean(x), 1)
)
```

## Interpretation

The synthetic dataset is designed to support realistic analytics workflows:

- assignment-family comparisons
- completion and missingness monitoring
- score-distribution analysis
- student-risk and engagement segmentation
- public-safe dashboard development

The report is not evidence about a real class or school. It is a reproducible
artifact demonstrating how private reference structure can be transformed into a
public-safe dataset for analysis and portfolio presentation.
